Construction of replicated designs based on Sobol' sequences

### 2016-05-12 crass/crass.tex
Most of section 2 reviewed

\begin{abstract}
\selectlanguage{english}
% Text of abstract in English
In the perspective of estimating main effects of model inputs, two approaches are studied to construct replicated designs based on Sobol' sequences.
{\it To cite this article: L.
Gilquin, L. A. J. Rugama, E. Arnaud, F. J. Hickernell, H. Monod, C. Prieur, C. R. Acad. Sci. Paris, Ser. I 340 (2005).}

@@

Digital nets and sequences were first introduced by Niederreiter \cite{Niederreiter} in the numerical integration framework to define good uniformly distributed points in $\cube$. They can also appear in the literature as digital $(t,m,s)$-nets and digital $(t,s)$-sequences, or simply $(t,m,s)$-nets and $(t,s)$-sequences. Sobol' and Niederreiter-Xing sequences are two examples whose explicit constructions are detailed in \cite{Niederreiter} and \cite{Sobol1}. %\cite{Niederreiter, Sobol, DickPhilischammer,DickKritzer,KuoJoe}.

@@

By increasing $m$, we increase the number of points of the $(t,m,s)$-net. If one considers the limit for $m\rightarrow\infty$, $(t,s)$-sequences are defined as:
\begin{definition}%[$(t,s)$-sequence]
For integers $s\geq 1$, $b\geq 2$, and $t\geq 0$, the sequence $\{\vx_i\}_{i\in\mathbb{N}_0}$ is a $(t,s)$-sequence in base $b$, if for every set $\mathcal{P}_{\ell,m}=\{\vx_i\}_{i=\ell b^m}^{(\ell+1)b^m-1}$ with $\ell\geq 0$ and $m\geq t$, $\mathcal{P}_{\ell,m}$ is a $(t,m,s)$-net in base $b$.
\end{definition}
\bigskip

All $(t,m,s)$-nets and $(t,s)$-sequences relevant for applications are based on a digital construction scheme over a finite field $\mathbb{F}_b:=\{0,\dots,b-1\}$ with $b$ a prime power number. Digital nets and sequences referred to these structures.

The construction scheme of a $s$-dimensional digital sequence requires the knowledge of the base $b$ and a set of $s$ full rank upper triangular matrices of infinite size called generating matrices. These generating matrices are recursively constructed through primitive polynomials and initial directional numbers (see \cite{Niederreiter} for further details). We detail below the construction scheme.

Denote by $C_1,\dots,C_s$ the $s$ generating matrices over $\mathbb{F}_b$ of a $s$-dimensional digital sequence. For $m>0$ and $j \in \{1,\dots,s\}$, we note $C_j^{m}$ the $m \times m$ upper left block of $C_j$. The first $b^m$ points $\{\vx_0,\dots,\vx_{b^m-1}\}$ of the digital sequence are constructed as follows:

For each $i=0,\dots,b^m-1$, the point $\vx_i = (x_{i,1},\dots,x_{i,s})^\intercal$ of the sequence is obtained dimension-wise by:
\begin{equation}
\label{dig.net.eq.}
(x_{i,j,1},x_{i,j,2},\dots)^\intercal = C_j^{m} \vi,\qquad j= 1,\dots,s\, ,
\end{equation}
where $x_{i,j} = \sum_{k \geq 1}x_{i,j,k}b^{-k}$ and $\vi = (i_{0},\dots,i_{m-1})^\intercal$ is the b-ary decomposition of $i$.
\bigskip

@@

The concept of digital sequences can be linked to the definition of replicated designs. To do so, we introduce the following notion:
\begin{definition}
Two digital sequences $\{\vx_i\}_{i\in\mathbb{N}_0}$ and $\{{\vx'}_i\}_{i\in\mathbb{N}_0}$ are said \emph{digitally replicated} of order $r$ if for all $m\geq 0$, $\{{\vx}_i\}_{i=0}^{b^m-1}$ and $\{{\vx'}_i\}_{i=0}^{b^m-1}$ are two replicated designs of order $r$.
\end{definition}

Sobol' sequences are digital sequences constructed in base $2$. These sequences are attractive for their easy implementation and optimized properties \cite{Kuo}. Sobol' sequences are particularly interesting for having full rank upper triangular generating matrices over $\mathbb{F}_2$. This property leads to the following lemma:
\begin{lemma}\label{Sobol_replicated}
All Sobol' sequences are digitally replicated of order 1.
\end{lemma}
\begin{pf*}{Proof.}
Consider any two s-dimensional Sobol' sequences generated respectively by the generating matrices $C_1,\dots,C_s$ and $C'_1,\dots,C'_s$. Since these matrices are upper triangular, to compute the first $2^m$ points one only needs the knowledge of the $m\times m$ upper left blocks $C_j^m$ and ${C'}_j^m$, $j \in \{1,\dots,s\}$. Because any two matrices $C^{m}_j$ and ${C'}^{m}_j$ are square and full rank, the products $C^{m}_j\vi$ and ${C'}^{m}_j\vi$ are one-to one and onto, for all $i=0,\dots,2^m-1$. Therefore, they generate the same point sets.
\end{pf*}

@@

\section{Iterative construction of replicated point sets}

We propose here two approaches to construct two replicated point sets $\mathcal{P}$ and $\mathcal{P}'$ based on Sobol' sequences. These two constructions are iterative that is the two points sets constructed possess the following structure:
$$\mathcal{P}= B_1 \cup B_2 \cup B_3 \dots $$
$$\mathcal{P}'= {B'}_1 \cup {B'}_2 \cup {B'}_3 \dots $$
where $B_l$ (resp. $B'_l$), $l\geq1$, is a new set of points added at step $l$ to refine $\mathcal{P}$ (resp. $\mathcal{P}'$). For each approach and at each step, $\mathcal{P}$ and $\mathcal{P'}$ are two replicated designs of order $1$.

@@

>>>>

\begin{abstract}
\selectlanguage{english}
% Text of abstract in English
In the perspective of estimating main effects of model inputs, two approaches are studied to construct replicated designs based on Sobol' sequences.
{\it To cite this article: L.
Gilquin, Ll. A. Jim\'enez Rugama, E. Arnaud, F. J. Hickernell, H. Monod, C. Prieur, C. R. Acad. Sci. Paris, Ser. I 340 (2005).}

@@

Digital nets and sequences were first introduced by Niederreiter \cite{Niederreiter} in the numerical integration framework to define good uniformly distributed points in $\cube$. They can also appear in the literature as digital $(t,m,s)$-nets and digital $(t,s)$-sequences, or simply $(t,m,s)$-nets and $(t,s)$-sequences. Sobol' and Niederreiter-Xing sequences are two examples of digital sequences detailed in \cite{Sobol1} and \cite{NiederreiterXing}.

@@

By increasing $m$, we increase the number of points of the $(t,m,s)$-net. In the limit case where $m\rightarrow\infty$, we can define $(t,s)$-sequences as:
\begin{definition}%[$(t,s)$-sequence]
For integers $s\geq 1$, $b\geq 2$, and $t\geq 0$, the sequence $\{\vx_i\}_{i\in\mathbb{N}_0}$ is a $(t,s)$-sequence in base $b$, if for every set $\mathcal{P}_{\ell,m}=\{\vx_i\}_{i=\ell b^m}^{(\ell+1)b^m-1}$ with $\ell\geq 0$ and $m\geq t$, $\mathcal{P}_{\ell,m}$ is a $(t,m,s)$-net in base $b$.
\end{definition}

Replicated design properties can apply to digital sequences. Therefore, we introduce the definition,
\begin{definition}
Two digital sequences $\{\vx_i\}_{i\in\mathbb{N}_0}$ and $\{{\vx'}_i\}_{i\in\mathbb{N}_0}$ are said \emph{digitally replicated} of order $r$ if for all $m\geq 0$, $\{{\vx}_i\}_{i=0}^{b^m-1}$ and $\{{\vx'}_i\}_{i=0}^{b^m-1}$ are two replicated designs of order $r$.
\end{definition}

$s$-dimensional Sobol' sequences are designed in base $b=2$ and can be computed using the \emph{generating matrices}, a set of $s$ full rank infinite dimensional upper triangular matrices defined over $\mathbb{F}_2:=\{0,1\}$. In \cite{Kuo}, Kuo and Joe suggest a particular choice for these matrices that optimize the 2 dimensional projection $t$-values.

We denote these matrices by $C_1,\dots,C_s$, and $C_1^m,\dots,C_s^m$ their upper left corner blocks of size $m\times m$. Although $C_1,\dots,C_s$ have infinite size, one only requires the knowledge of $C_1^m,\dots,C_s^m$ to construct the first $2^m$ points.

For each $i=0,\dots,2^m-1$, the point $\vx_i = (x_{i,1},\dots,x_{i,s})^\intercal$ of the sequence is obtained dimension-wise by:
\begin{equation}
\label{dig.net.eq.}
(x_{i,j,1},x_{i,j,2},\dots)^\intercal = C_j^{m} \vi,\qquad j= 1,\dots,s\, ,
\end{equation}
where $x_{i,j} = \sum_{k \geq 1}x_{i,j,k}2^{-k}$ and $\vi = (i_{0},\dots,i_{m-1})^\intercal$ the binary decomposition of $i$. Below we show how to compute $x_{11,1}=0.8125$ and $x_{9,2}=0.4375$

\begin{equation*}
\begin{pmatrix}
x_{11,1,1} \\
x_{11,1,2} \\
x_{11,1,3} \\
x_{11,1,4}
\end{pmatrix} =
\begin{pmatrix}
1 \\
1 \\
0 \\
1
\end{pmatrix} = 
\overbrace{
\begin{pmatrix}
1 & 0 & 0 & 0 \\
0 & 1 & 0 & 0 \\
0 & 0 & 1 & 0 \\
0 & 0 & 0 & 1
\end{pmatrix}}^{C_1^4}
\overbrace{
\begin{pmatrix}
1 \\
1 \\
0 \\
1
\end{pmatrix}}^{\bvec{11}},\qquad
\begin{pmatrix}
x_{9,2,1} \\
x_{9,2,2} \\
x_{9,2,3} \\
x_{9,2,4}
\end{pmatrix} =
\begin{pmatrix}
0 \\
1 \\
1 \\
1
\end{pmatrix} = 
\overbrace{\begin{pmatrix}
1 & 1 & 1 & 1 \\
0 & 1 & 0 & 1 \\
0 & 0 & 1 & 1 \\
0 & 0 & 0 & 1
\end{pmatrix}}^{C_2^4}
\overbrace{\begin{pmatrix}
1 \\
0 \\
0 \\
1
\end{pmatrix}}^{\bvec{9}}
\end{equation*}
\bigskip

@@

\begin{lemma}\label{Sobol_replicated}
All Sobol' sequences are digitally replicated of order 1.
\end{lemma}
\begin{pf*}{Proof.}
Consider any two $s$-dimensional Sobol' sequences generated by the generating matrices $C_1,\dots,C_s$ and $C'_1,\dots,C'_s$ respectively. Since these matrices are upper triangular, to compute the first $2^m$ points one only needs the knowledge of the $m\times m$ upper left blocks $C_j^m$ and ${C'}_j^m$, $j \in \{1,\dots,s\}$. Because any two matrices $C^{m}_j$ and ${C'}^{m}_j$ are square and full rank, the operations $C^{m}_j\vi$ and ${C'}^{m}_j\vi$ are one-to one and onto, for all $i=0,\dots,2^m-1$. Therefore, they generate the same point sets.
\end{pf*}

@@

\section{Iterative construction of replicated point sets}
SCRAMBLING
We propose here two approaches to construct two replicated point sets $\mathcal{P}$ and $\mathcal{P}'$ based on Sobol' sequences. These two constructions are iterative that is the two points sets constructed possess the following structure:
$$\mathcal{P}= B_1 \cup B_2 \cup B_3 \dots $$
$$\mathcal{P}'= {B'}_1 \cup {B'}_2 \cup {B'}_3 \dots $$
where $B_l$ (resp. $B'_l$), $l\geq1$, is a new set of points added at step $l$ to refine $\mathcal{P}$ (resp. $\mathcal{P}'$). For each approach and at each step, $\mathcal{P}$ and $\mathcal{P'}$ are two replicated designs of order $1$.

@@

\bibitem{NiederreiterXing} H. Niederreiter, C. Xing, Low-Discrepancy Sequences and Global Function Fields with Many Rational Places, \emph{Finite Fields and Their Applications} 2(3), 1996, pp. 241--273.
